---
title: "Homework 2"
author: "Jiaming Chen"
date: "12/19/2017"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## 1

calculate n

n = 13.09777

```{r}
power.t.test(delta = 1, sd = 1, power = 0.8, type = "two.sample", alternative = "one.sided")
```

How n change along with $\Delta$?

n decrease when $\Delta$ increase.

```{r}
power.t.test(delta = 2, sd = 1, power = 0.8, type = "two.sample", alternative = "one.sided")
power.t.test(delta = 3, sd = 1, power = 0.8, type = "two.sample", alternative = "one.sided")
power.t.test(delta = 4, sd = 1, power = 0.8, type = "two.sample", alternative = "one.sided")

```

How n change along with $\sigma^2_A$?

n increase when $\sigma^2_A$ increase.


```{r}
power.t.test(delta = 1, sd = 0.7, power = 0.8, type = "two.sample", alternative = "one.sided")
power.t.test(delta = 1, sd = 1, power = 0.8, type = "two.sample", alternative = "one.sided")
power.t.test(delta = 1, sd = 2, power = 0.8, type = "two.sample", alternative = "one.sided")
```


## 2

We should use Chi-Square test, because we're checking if a random number follows a certain distribution.

N=101

$~~$|0 six| 1 six| 2 six| 3 six
------- | ---------- | ---------- | ---------- | ----------
expected|58.58|35.35|7.07|0.4646
observed|48|35|15|3

$\chi^2=1.91+0.0035+8.89+13.84=24.6435$

$df = (2-1)\times(4-1)=3$

p-value is 1.833077e-05
```{r}
1-pchisq(24.6435,3)
```

## 3

Feature engineering:

The new features can be generated by combining existing features or splitting existing features. For example, for the categorical features with too many levels, we can combine some levels to form a new categorical feature with less levels. We can also combine correlated features to avoid colinearity.
