---
title: "Untitled"
author: "Jiaming Chen"
date: "12/12/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)    # my usual tools, ggplot2, dplyr
```

# Correlation and Causation




# WAHT IS HYPOTHESIS TEST?

## A/B Testing

Define the causal relationship to be explored

Define metric

Design randomized experiments (A/B test)

Collect data and conduct ~hypothesis testing~

Draw conclusion



> Null hypothesis, H0 , is usually the hypothesis that sample observations result purely from chance.
> Alternative hypothesis, H1 or Ha, is the hypothesis that sample observations are influenced by some non-random cause

# TEST ABOUT MEAN

## Type I and Type II error

Type I error: H0 is true but we reject it $\alpha$
Type II error: H0 is false but we accept it $\beta$

Power + $\beta$ = 1

reduce error rate --> increase sample size and decrease variance 

"LIFT" -- when we want to detect a small amount of LIFT, we need to decide how many samples we have to reach a certain Power (let's say 80%)


T test is more conservative than Z test

Pooled variance -- a weighted sum of variance from x1 and variance from x2
Pooled --> student t test: equal variance, normality, independence
unpooled --> welch test

# TEST ABOUT VARIANCE

Chi square test (always one sided)

It's used for:

> test population variance
> goodness of fit (test), does a categorical variable has a certain distribution?
__example: test if a coin is a fair coin or not. (observed and expected)__
> null and alternative hypothesis
> compare two population variance
> compare two categorical variable to see if they have the same distribution or not


```{r}
loanT <- loan

num.value <- sapply(loan, function(x) {return(length(unique(x)))})

which(num.value == 1)
which(num.value == dim(loan)[1])
```

## Lending club

if we take Interest Rate as the response, we cannot include Loan Payment feature (because we don't have them when we decide int_rate)


```{r}
summary(loanT$dti_joint)
loanT$dti <- ifelse(!is.na(loanT$dti_joint), loanT$dti_joint, loanT$dti)
loanT$annual_inc <- ifelse(!is.na(loanT$annual_inc_joint), loanT$annual_inc_joint, loanT$annual_inc)


#loanT$home_ownership <- ifelse()
```


